Using Hybrid Machine Learning for Fundamental Stellar Parameter Inference

### 2024-04-16 paper/paper2_neo/paper.tex
Write multiprocessing full grid inference

Naturally, scientists want to extract as much information as possible from
a star's spectra to get back the properties that lie at the root of it all. 
However, this has proven to be no easy task. Spectroscopic surveys such as 
APOGEE develop their own in-house pipelines to extract stellar parameters
from spectra (ASPCAP in the case of APOGEE), and these pipelines are fairly
effective. However, they all work on the core assumption that the data to be 
analyzed is a list of pixels. In addition, the pipelines that surveys 
develop are closed-source and usually limited to the scope of the survey 
itself, causing a sort of fragmentation. This motivates the development of a 
more universal, open-source tool that can do what spectral analysis pipelines 
do, but in an alternative way.

Multiple efforts have been made in this direction, treating spectra in 
different ways. Some take the wavelength and flux values and construct
neural networks to create this mapping between stellar parameters and the
observed spectra, such as \texttt{IN-SYNC}. Others take a more 
physics-informed approach by decomposing spectra into some eigenbasis and 
treating it as a regression problem, such as \texttt{starfish}.
However, \texttt{blase} takes a different approach. It treats spectra not as 
a set of pixels or a set of eigenbasis coefficients but as a set of 
spectral lines, specifically Voigt profiles. There is no metric of this 
approach being better or worse than the others mentioned so far, but it is 
a different paradigm and is worth exploring as a semi-empirical tool for
spectral inference. This study aims to take advantage of \texttt{blase}'s
ability to change a spectrum's basis from flux versus wavelength into
spectral line versus spectral line properties, and use that to develop a
robust tool for stellar parameter inference.

@@

\begin{table}[h]
    \hspace*{-1.1cm}\begin{tabular}{|c|c|c|}
        \hline
        \multicolumn{3}{|c|}{\textbf{PHOENIX Grid Subset}}\\
        \hline\hline
        Parameter & Symbol & Range\\
        \hline
        Alpha Element Abundance & $\alpha$ & [0] dex\\
        Iron Abundance & [Fe/H] & [-0.5, 0] dex\\
        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K\\
        Surface Gravity & $\log(g)$ & [2, 6]\\
        Wavelength & $\lambda$ & [8038, 12849] \AA\\
        \hline
    \end{tabular}
    \caption{The subset of the PHOENIX grid used in this study. These limits 
    were imposed to reduce the computational cost of the algorithms and to 
    ensure a rectilinear parameter space in order to work with the chosen
    interpolator backend. The wavelength limits in particular roughly line 
    up with that of the Habitable Zone Planet Finder (HPF) spectrograph.}
\end{table}

\subsection{Preprocessing with \texttt{gollum}}
First, the PHOENIX subset was accessed directly from the PHOENIX website
using \texttt{gollum}'s download option, which uses the website's FTP server
to download spectra without using any disk space. This came at the cost of
longer load times due to internet connection being significantly more of a
bottleneck than disk I/O, but was acceptable as this was a one-time operation.

The spectra were then put through a three-step preprocessing pipeline.
\begin{enumerate}
    \item Blackbody Division: Since the $T_{\mathrm{eff}}$ of each spectrum 
    is known, the according blackbody spectrum was divided out.
    \item Percentile Normalization: The spectra were normalized by dividing
    them by their 99th percentile.
    \item Continuum Normalization: The spectra were further normalized by 
    dividing them by a \nth{5} order polynomial continuum.
\end{enumerate}

>>>>

Scientists want to extract as much information as possible from
a star's spectra to get back the properties that lie at the root of it all. 
However, this has proven to be no easy task. Spectroscopic surveys such as 
APOGEE develop their own in-house pipelines to extract stellar parameters
from spectra (ASPCAP in the case of APOGEE), and these pipelines are fairly
effective. However, they all work on the core assumption that the data to be 
analyzed is a list of pixels. In addition, the pipelines that surveys 
develop are typically closed-source and usually limited to the scope of the 
survey itself, causing a sort of fragmentation. This motivates the 
development of a more universal, open-source tool that can do what spectral 
analysis pipelines do, but in an alternative way.

Multiple efforts have been made in this direction, treating spectra in 
different ways. The standard practice is to treat the wavelength and flux as 
simply two arrays and implement any from a myriad of different algorithms to 
inference fundamental stellar parameters. Others take a more physics-informed 
approach by decomposing spectra into some eigenbasis and treating it as a 
regression problem, such as \texttt{starfish}. However, \texttt{blase} takes 
a different approach. It treats spectra not as a set of pixels or a set of 
eigenbasis coefficients but as a set of spectral lines, specifically Voigt 
profiles. There is no metric of this approach being better or worse than the 
others mentioned so far, but it is a different paradigm and is worth 
exploring as a semi-empirical tool. This study aims to take advantage of 
\texttt{blase}'s ability to change a spectrum's basis from flux versus 
wavelength into spectral line versus spectral line properties, and use that 
to develop a robust tool for stellar parameter inference.

@@

\begin{table}[h]
    \hspace*{-1.35cm}\begin{tabular}{|c|c|c|}
        \hline
        \multicolumn{3}{|c|}{\textbf{PHOENIX Grid Subset}}\\
        \hline\hline
        Parameter & Symbol & Range\\
        \hline
        Alpha Element Abundance & $\alpha$ & [0] dex\\
        Iron Abundance & [Fe/H] & [-0.5, 0] dex\\
        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K\\
        Surface Gravity & $\log(g)$ & [2, 6]\\
        Wavelength & $\lambda$ & [8038, 12849] \AA\\
        \hline
    \end{tabular}
    \caption{The subset of the PHOENIX grid used in this study. These limits 
    were imposed to reduce the computational cost of the algorithms and to 
    ensure a rectilinear parameter space in order to work with \texttt{scipy}'s
    \texttt{RegularGridInterpolator}. The wavelength limits in particular 
    roughly line up with that of the Habitable Zone Planet Finder (HPF) 
    spectrograph.}
\end{table}

\subsection{Preprocessing with \texttt{gollum}}
First, the PHOENIX subset was accessed directly from the PHOENIX website
using \texttt{gollum}'s download option. The spectra were then put through a 
three-step preprocessing pipeline.
\begin{enumerate}
    \item Blackbody Division: Since the $T_{\mathrm{eff}}$ of each spectrum 
    is known, the according blackbody spectrum was divided out.
    \item Percentile Normalization: The spectra were normalized by dividing
    them by their 99th percentile.
    \item Continuum Normalization: The spectra were further normalized by 
    dividing them by a \nth{5} order polynomial continuum.
\end{enumerate}
